Analytic Light Curves in Reflected Light: \\% Phase Curves, Occultations, and Non-Lambertian Scattering

### 2020-05-08 tex/ms.tex
added figure

% Title
\title{Analytic Occultation Light Curves in Reflected Light}

@@

Our goal in this paper is to derive analytic expressions for the
flux received by a distant observer from a non-uniform Lambert sphere
illuminated by a monochromatic source that may or may not be
occulted by a (possibly different) spherical body. This applies, for example,
to the
case of exoplanet phase curves, exoplanet secondary
eclipse (occultation) light curves, or exoplanet-moon and exoplanet-exoplanet
occultations, all seen in reflected light. We explicitly assume the reflecting
body is Lambertian, i.e., it scatters light isotropically, although with an
intensity that varies across the surface according to a spatially-dependent
albedo $A$. Throughout this paper, we will take $A$ to mean the \emph{spherical}
albedo, the fraction of power incident on a body at a given wavelength
that is scattered back out to space (in all directions).
Note that the spherical albedo
is closely related to the Bond albedo: the Bond albedo is the stellar
flux-weighted integral of $A(\lambda)$ over all wavelengths $\lambda$
\citep[see, e.g.,][]{Seager2010}.

As in \citet{Luger2019}, we compute fluxes by first expanding the surface
in terms of spherical harmonics.
While in \citet{Luger2019} we expanded the emissivity of the surface, here
we instead expand the spherical albedo $A$.
Specifically, if $\bvec{y}$ is the vector
of spherical harmonic coefficients describing the albedo anywhere on
the surface and
$\by$ is the spherical harmonics basis (Equation~\ref{eq:by}),
the albedo $A$ at a point $(x, y)$ on the sky-projected disk of the body
is given by the dot product
%
\begin{align}
    \label{eq:albedo}
    A(x, y) = \by^\top (x, y) \, \bvec{y}
    \quad.
\end{align}
%
The flux measured from this body is proportional to the surface integral
over the projected disk
of the albedo $A$ times the illumination profile $\mathcal{I}$ of the surface,
given by Lambert's law as
%
\begin{align}
    \label{eq:LambertsLaw}
    \mathcal{I}(\vartheta_i) = \mathcal{I}_0 \, \text{max}\big( 0, \cos\vartheta_i \big)
    \quad,
\end{align}
%
where $\vartheta_i$ is the angle between the incident radiation and the surface
normal and $\mathcal{I}_0$ is the peak illumination. We show in
Appendix~\ref{sec:adapting-starry} that in the case of, say, a planet
illuminated by its host star,
%
\begin{align}
    \mathcal{I}_0 = \frac{f_s}{\pi r_s^2}
\end{align}
%
where $r_s$ is the distance between the planet and the star (in units
of the planet's radius) and $f_s$ is the stellar flux measured at the observer
(in arbitrary units). Following the convention in \citet{Luger2019}, we assume
throughout this paper that $f_s = 1$, so all fluxes are defined as a fraction of
the flux of the illumination source.

@@

>>>>

% Title
\title{%
    Analytic Light Curves in Reflected Light:%
    \\%
    Phase Curves, Occultations, and Non-Lambertian Scattering
}

@@

Our goal in this paper is to derive analytic expressions for the
flux received by a distant observer from a sphere of non-uniform albedo
illuminated by a monochromatic source that may or may not be
occulted by a (possibly different) spherical body. This applies, for example,
to the case of planetary phase curves, secondary
eclipse (occultation) light curves, and moon-moon, planet-moon, and
planet-planet occultations (in the Solar System or not),
all seen in reflected light. We derive all expressions in
the limit that the reflecting
body is Lambertian, i.e., it scatters light isotropically, but we relax
this assumption in later sections.
We model the general case of an
intensity that varies across the surface of the body according to a
spatially-dependent
albedo $A$. Throughout this paper, we will take $A$ to mean the \emph{spherical}
albedo, the fraction of power incident on a body at a given wavelength
that is scattered back out to space (in all directions).
Note that the spherical albedo
is closely related to the Bond albedo: the Bond albedo is the stellar
flux-weighted integral of $A(\lambda)$ over all wavelengths $\lambda$
\citep[see, e.g.,][]{Seager2010}.

As in \citet{Luger2019}, we compute fluxes by first expanding the surface
in terms of spherical harmonics.
While in \citet{Luger2019} we expanded the emissivity of the surface, here
we instead expand the spherical albedo $A$.
Specifically, if $\bvec{y}$ is the vector
of spherical harmonic coefficients describing the albedo anywhere on
the surface and
$\by$ is the spherical harmonics basis (Equation~\ref{eq:by}),
the albedo $A$ at a point $(x, y)$ on the sky-projected disk of the body
is given by the dot product
%
\begin{align}
    \label{eq:albedo}
    A(x, y) = \by^\top (x, y) \, \bvec{y}
    \quad.
\end{align}
%
The flux measured from this body is proportional to the surface integral
over the projected disk
of the albedo $A$ times the illumination profile $\mathcal{I}$ of the surface,
given by Lambert's law as
%
\begin{align}
    \label{eq:LambertsLaw}
    \mathcal{I}(\vartheta_i) = \mathcal{I}_0 \, \text{max}\big( 0, \cos\vartheta_i \big)
    \quad,
\end{align}
%
where $\vartheta_i$ is the angle between the incident radiation and the surface
normal and $\mathcal{I}_0$ is the peak illumination. We show in
Appendix~\ref{sec:adapting-starry} that in the case of, say, a planet
illuminated by its host star,
%
\begin{align}
    \mathcal{I}_0 = \frac{f_s}{\pi r_s^2}
\end{align}
%
where $r_s$ is the distance between the planet and the star (in units
of the planet's radius) and $f_s$ is the stellar flux measured at the observer
(in arbitrary units). Following the convention in \citet{Luger2019}, we assume
throughout this paper that $f_s = 1$, so all fluxes are defined as a fraction of
the flux of the illumination source at the observer.

@@

\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/inference.pdf}
        \oscaption{inference}{%
            Example of an inference problem for a thermal phase curve
            (center left) and a reflected phase curve (center right).
            In both cases,
            a mock phase curve is generated from an $l=20$ expansion
            of the cloudless Earth (top) with 1,000 evenly spaced points
            over the course of one year and an extremely small
            photometric uncertainty of 1 ppm. The observer sits along the
            ecliptic and the obliquity of the Earth is set to $23.5^\circ$.
            Data is shown as the black points, and the maximum likelihood
            \starry model is shown in blue. While both models fit the
            data equally well, the same is not true of the inferred
            surface maps (bottom row): only in the reflected case are the
            continental outlines recovered.
            The thermal phase curve problem is
            extremely ill-conditioned, but the analogous problem in reflected
            light is much better posed.
            \label{fig:inference}
        }
    \end{centering}
\end{figure}
